Predicting the spread and persistence of genetically modified dominant sterile male mosquitoes

### 2021-07-27 inst/manuscript/Draft/response_to_reviewers.tex

\documentclass[12pt]{paper}

\usepackage{latexsym,amssymb,amsfonts,amsmath}
\usepackage[pdftex]{graphicx}
% or use the graphicx package for more complicated commands
\usepackage{graphicx, graphics}
\usepackage[round]{natbib}
\usepackage{blindtext,tikz}
\usetikzlibrary{calc}
\usepackage{datetime}
\usepackage[obeyspaces]{url}
\usepackage{amsthm, amsmath, amssymb}
\usepackage{geometry}
\geometry{margin = 60pt}
\usepackage{setspace}
\doublespacing
%% The lineno packages adds line numbers. Start line numbering with

%%% Put your definitions there:
\title{List of changes: addressing the comments}

\begin{document}

\maketitle

%\linenumbers
We would like to thank again the reviewers for this last round of comments. We addressed this time the remarks on reviewer-by-reviewer basis. \\

\section*{Reviewer A:}

\emph{
\footnotesize
\begin{itemize}
\item  \\
{\color{red} - . }
\item I believe that one of the reviewers asked for more information regarding the model by Szpiro et al. (2010) on which this model builds on, but the authors have not provided any background information. I think this should be included. In particular, it would be worthwhile to understand why is the paper by Szpiro et al. (2010) not appropriate for this data. Why is there a need for a mixture model?\\
\end{itemize}
}
~\\
~\\

\newpage

\section*{Reviewer B:}
\emph{
\footnotesize
\begin{itemize}
\item  \\
{\color{red} }
\end{itemize} 
}
~\\
~\\
\end{document}

%%
%% End of file `elsarticle-template-1a-num.tex'.

### 2021-07-27 inst/manuscript/Draft/bmc_article_revision_v1.tex
initialise resposne to reveiwers

Our objective is to predict the spread and persistence of the genetically modified mosquitoes ahead of a proposed field release. These predictions are an important component of an independent risk assessment conducted prior to the field release \cite{Hayes2015a}, and could also help to inform the design of a post-release monitoring strategy. Our approach implements the Bayesian paradigm; we elicit the prior distribution of model parameters from relevant experts, for both the wild type mosquito (WT) and its DSM counterpart. We obtain the posterior model parameter values for the WT parameters using the results of mark-release-recapture (MRR) experiments conducted by \cite{Epopa2017}. We then use Bayesian model averaging to predict the spread and persistence of DSM mosquitoes.

@@

\section*{Competing interests}
  The authors declare that they have no competing interests.

\section*{Author's contributions}
AI carried out the original draft preparation, proposed the statistical model and ran the statistical analyses; SF wrote the code for the PDE model. KH and GH participated in expert elicitation. All authors participated on drafting and revision of the manuscript. All authors gave final approval for publication and 
agree to be held accountable for the work performed herein.

\section*{Acknowledgements}
 This study would not be possible without the contribution of the domain experts - Heather Ferguson, Tovi Lehmann and Fred Gould; The authors would also like to acknowledge the numerous discussions and feedback received from Ron Thresher, Uffe Høgsbro Thygesen and Peter Caley. The authors would like to thank Jess Ford, Jeff Dambacher and Anders Gonçalves da Silva for helping facilitate the expert elicitation sessions. This analysis was funded by the Foundation for the National Institutes of Health and CSIRO Health and Biosecurity. The authors would also like to acknowledge CSIRO Data61 and Target Malaria for their help and support.

\section*{Ethics}
All elicitations were conducted with the informed consent of independent experts under ethics approval from the CSIRO Social Science Human Research Ethics Committee (reference LR06/2014). The mark release capture data was kindly provided by the Institut de Recherche en Sciences de la Sant\'{e}, Bobo-Dioulasso, Burkina Faso, and was collected with approval from the local institutional ethics committee (Centre Muraz Institutional Ethics Committee), reference number 009–2012/CE-CM.

\section*{Availability of Data and Materials}

>>>>

Our objective is to predict the spread and persistence of the genetically modified mosquitoes ahead of a proposed field release. These predictions are an important component of an independent risk assessment conducted prior to the field release \cite{Hayes2015a} that assessed the risk of vecotr-borne disease transmission \cite{Hosack2020}, and could also help to inform the design of a post-release monitoring strategy. Our approach implements the Bayesian paradigm; we elicit the prior distribution of model parameters from relevant experts, for both the wild type mosquito (WT) and its DSM counterpart. We obtain the posterior model parameter values for the WT parameters using the results of mark-release-recapture (MRR) experiments conducted by \cite{Epopa2017}. We then use Bayesian model averaging to predict the spread and persistence of DSM mosquitoes.

@@

\section*{Declaration}

\subsection*{Acknowledgements}
 This study would not be possible without the contribution of the domain experts - Heather Ferguson, Tovi Lehmann and Fred Gould; The authors would also like to acknowledge the numerous discussions and feedback received from Ron Thresher, Uffe Høgsbro Thygesen and Peter Caley. The authors would like to thank Jess Ford, Jeff Dambacher and Anders Gonçalves da Silva for helping facilitate the expert elicitation sessions. This analysis was funded by the Foundation for the National Institutes of Health and CSIRO Health and Biosecurity. The authors would also like to acknowledge CSIRO Data61 and Target Malaria for their help and support.

\subsection*{Author's contributions}
AI carried out the original draft preparation, proposed the statistical model and ran the statistical analyses; SF wrote the code for the PDE model. KH and GH participated in expert elicitation. All authors participated on drafting and revision of the manuscript. All authors gave final approval for publication and 
agree to be held accountable for the work performed herein.

\subsection*{Funding}
Not applicable.

\subsection*{Consent for publication}
Not applicable.

\subsection*{Competing interests}
The authors declare that they have no competing interests.

\subsection*{Ethics approval and consent to participate}
All elicitations were conducted with the informed consent of independent experts under ethics approval from the CSIRO Social Science Human Research Ethics Committee (reference LR06/2014). The mark release capture data was kindly provided by the Institut de Recherche en Sciences de la Sant\'{e}, Bobo-Dioulasso, Burkina Faso, and was collected with approval from the local institutional ethics committee (Centre Muraz Institutional Ethics Committee), reference number 009–2012/CE-CM.

\subsection*{Availability of data and material}
